---
title: "Project2"
author: "Hiba Hajali"
date: "20/03/2022"
output:
  pdf_document:
    latex_engine: pdflatex
    number_sections: yes
    keep_tex: true
  html_document:
    df_print: paged
fig_caption: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(janitor)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(gridExtra)
library(rcompanion)
library(knitr)
library(kableExtra)
library(pROC)
library(caret)
library(huxtable)
library(grid)
```

# Introduction 

Coffee is one of the most popular beverages worldwide and has a vast market. Research on coffee quality can help coffee farmers understand the quality of the coffee they grow to make more accurate market planning. The researchers obtained data containing features of coffee and its production from the Coffee Quality Institute, a coffee research institute. They used this data to analyze the impact of these coffee features (such as acidity) on coffee quality scores. In the following sections, the researchers will use the Generalized Linear Model to model the Qualityclass variables, obtain the optimal model by comparison, and analyze each variable to determine its impact on coffee quality.
```{r}
##import data 
data <- read.csv("dataset8.csv")
```
# Explanatory Analysis 

The numbers of the missing values in each column:


```{r}

##check for missing values using sapply
sapply(data, function(x) sum(is.na(x)))

```

The data after we remove the missing values:

```{r}
##removing missing values
data <- na.omit(data)
glimpse(data)

##replace good/poor with the value 1 and 0
#data$Qualityclass <- ifelse(data$Qualityclass == "Good", 1, 0)

##assign the value to keep in variable
#Qualityclass <- data$Qualityclass
#country_of_origin <- data$country_of_origin
```



The number of unique values in country of origin: 

```{r}
##check for number of unique values in country of origin and harvest year using sapply
length(unique(data$country_of_origin))

```

The number of unique values in harvest year:
```{r}
length(unique(data$harvested))
```

The correlation between the quantitative variables:

```{r plots, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE}
##correlation between variables. 
pair_cor <- data %>% 
    select(Qualityclass,
           harvested,
           altitude_mean_meters,
           aroma,
           flavor,
           acidity,
           category_two_defects) 
pair_cor$Qualityclass <- as.factor(pair_cor$Qualityclass)


#Quality class vs aroma.
plot_1  <-  ggplot(data = pair_cor, 
                   aes(x = Qualityclass, 
                       y = aroma, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality class vs aroma.", y = "Aroma") +
            theme(legend.position = "none")

ggpairs(pair_cor) +
    theme(text = element_text(size=5))

```

Box plots showing the distribution of the quantitative variables  

```{r boxplots, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
plot__1  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = aroma, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Aroma") +
            theme(legend.position = "none")

plot_2  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = flavor, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Flavor") +
            theme(legend.position = "none")

plot_3  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = acidity, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Acidity") +
            theme(legend.position = "none")

plot_4  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = category_two_defects, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Category Two Defects") +
            theme(legend.position = "none")

plot_5  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = altitude_mean_meters, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Altitude mean (meters)") +
            theme(legend.position = "none")



grid.arrange(plot__1,plot_2,plot_3,nrow=1)
grid.arrange(plot_4, plot_5, nrow=1)
```


## bar charts:

```{r barp, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
#bar chart showing the quality class in each country 
plot_6  <-  ggplot(data = data, 
                   aes(x = country_of_origin, 
                       group = Qualityclass)) +
            geom_bar(aes(y = ..prop..,fill=  Qualityclass),
                     stat = "count", position = "dodge") + 
  theme(axis.text.x=element_text(angle=45, hjust=1), text=element_text(size=6))

#bar chart showing the quality class in each harvest year 
plot_7  <-  ggplot(data = data, 
                   aes(x = harvested, 
                       group = Qualityclass)) +
            geom_bar(aes(y = ..prop..,fill=  Qualityclass),
                     stat = "count", position = "dodge")
grid.arrange(plot_6, plot_7, ncol=1)


plot_8  <-  ggplot(data = data, 
                   aes(x = ..prop.., 
                       group = Qualityclass)) +
            geom_bar(aes(y = country_of_origin,fill=  Qualityclass),
                     stat = "count", position = "dodge") + 
  theme(axis.text.y =element_text(angle=0, hjust=1), text=element_text(size=10))
plot_8

plot_9  <-  ggplot(data = data, 
                   aes(x = country_of_origin, 
                       group = Qualityclass)) +
            geom_bar(aes(y = ..prop..,fill=  Qualityclass),
                     stat = "count", position = "dodge") + 
  theme(axis.text.x=element_text(angle=90, hjust=1))
plot_9
```

The percentages:

Table showing the percentage of the quality classes for each country

```{r,echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE, fig.align = "center",fig.pos = 'H',out.width = '68%'}
#table showing the percentage of the quality classes for each country  
data %>%
    tabyl(country_of_origin, Qualityclass) %>%
    adorn_percentages() %>%
    adorn_pct_formatting() %>%
    adorn_ns() %>% 
  kable(align='c',
   caption = "The Proportion of Quality Classs in Different Country") %>%
  kable_styling( latex_options="HOLD_position")

```

Table showing the percentage of the quality classes for each harvest year:

```{r, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE, fig.align = "center",  fig.pos = 'H', out.width = '68%'}
#table showing the percentage of the quality classes for each harvest year   
data %>%
    tabyl(harvested, Qualityclass) %>%
    adorn_percentages() %>%
    adorn_pct_formatting() %>%
    adorn_ns() %>% 
  kable(align='c',
   caption = "The Proportion of Quality Classs in Different Harvested Year") %>%
  kable_styling( latex_options="HOLD_position")
```


# Formal Anaylsis 
## Multilinearility Checking
```{r multilinearility???, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
multico.variables <- data %>% 
  select(aroma,
         flavor,
         acidity,
         category_two_defects,
         altitude_mean_meters,
         harvested) %>% 
  cor()

kappa(multico.variables, exact=TRUE)
##kappa value is 21.31, that is lower than 100. in general, it is suggest that the multilinearlity in our variables is quite small. 

```


Model 1:

$$ln\left(\frac{p_{Poor}}{1-p_{Poor}}\right) = \alpha + \beta_1 \cdot \textrm{Country} + \beta_2 \cdot \textrm{Aroma}+ \beta_3 \cdot \textrm{Flavor} + \beta_4 \cdot \textrm{Acidity} + \beta_5 \cdot \textrm{Category Two Defects}+\beta_6 \cdot \textrm{Harvested}+\beta_7 \cdot \textrm{Altitude}$$


```{r eval=TRUE, echo=FALSE, fig.align="center", fig.pos='H', message=FALSE, ,warning=FALSE, out.width='68%'}
#setting the quality class as factor 
data$Qualityclass <- as.factor(data$Qualityclass)

model_1 <- glm(formula = Qualityclass ~ .,family = binomial(link = "logit"),
               data = data)

```

Model 2:
$$ln\left(\frac{p_{Poor}}{1-p_{Poor}}\right) = \alpha + \beta_1 \cdot \textrm{Country} + \beta_2 \cdot \textrm{Aroma}+ \beta_3 \cdot \textrm{Flavor} + \beta_4 \cdot \textrm{Acidity} + \beta_5 \cdot \textrm{Category Two Defects}+\beta_6 \cdot \textrm{Harvested}$$


```{r, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
model_2 <- glm(formula = Qualityclass ~ country_of_origin +aroma+flavor+acidity+category_two_defects+harvested,family = binomial(link = "logit"),
               data = data)
```


Model 3:
$$ln\left(\frac{p_{poor}}{1-p_{poor}}\right) = \alpha + \beta_1 \cdot \textrm{Country of origin} + \beta_2 \cdot \textrm{aroma}+ \beta_3 \cdot \textrm{flavor} + \beta_4 \cdot \textrm{acidity} + \beta_5 \cdot \textrm{category two defects}$$


```{r model 3, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE,fig.align = "center",fig.pos = 'H', out.width = '68%'}
model_3 <- glm(formula = Qualityclass ~ country_of_origin +aroma+flavor+acidity+category_two_defects,family = binomial(link = "logit"),
               data = data)
```

Model 4:
$$ln\left(\frac{p_{Poor}}{1-p_{Poor}}\right) = \alpha + \beta_1 \cdot \textrm{Aroma}+ \beta_2 \cdot \textrm{Flavor} + \beta_3 \cdot \textrm{Acidity} + \beta_4 \cdot \textrm{Category Two Defects}$$

 
```{r model 4, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
model_4 <- glm(formula = Qualityclass ~ aroma+flavor+acidity+category_two_defects,family = binomial(link = "logit"),
               data = data)
```


Model 5:
$$ln\left(\frac{p_{Poor}}{1-p_{Poor}}\right) = \alpha  + \beta_1 \cdot \textrm{Aroma}+ \beta_2 \cdot \textrm{Flavor} + \beta_3 \cdot \textrm{Acidity}$$


```{r model 5, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
#best model
model_5 <- glm(formula = Qualityclass ~ aroma+flavor+acidity ,family = binomial(link = "logit"),
               data = data)
```


```{r baseline checking}
levels(data$Qualityclass)##check the baseline and contributor for our response variables. 
# the first returned result is our baseline and the second returned is the contributor. 
# it is means our model will estimates the change of the coffee bean with poor quality class according to the response baseline, which is good quality. While, this change is reflect the log-odds scale.
```


## Models comparison and Selection:




```{r model comparison, warning=FALSE, message=FALSE, fig.align = "center", fig.pos = 'H', out.height=12, out.width=6}
#compare 5 models
compareGLM(model_1, model_2, model_3, model_4,model_5) %>% 
  kable(digits=3,align='c',
   caption = "The Result of Model comparison") %>%
  kable_styling(font_size=8, latex_options="HOLD_position")

```




\newpage 

## log Odds: 
```{r odds plot, warning=FALSE, message=FALSE,out.width = '68%', fig.align = "center",  fig.pos = 'H'}
#Odds Plot
plot_model(model_5, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Poor instructor)", show.p = FALSE)

```



## Confidence Intervals:
```{r , warning=FALSE, message=FALSE, fig.align = "center",  fig.pos = 'H'}
#Confidence intervals
confint(model_5) %>% 
  kable(caption = '\\label{tab:CI5} Confidence Intervals for log odds in Model 5') %>% 
  kable_styling(font_size = 10,latex_options="HOLD_position")
```


## The Probability Plot: 
```{r probability plot,warning=FALSE, message=FALSE, fig.align = "center",  fig.pos = 'H' }
log.odds.est <- data %>%
  mutate(logodds.poor = predict(model_5))

plot_10 <- plot_model(model_5, type = "pred", title = "Probability of Poor Quality Class",
           axis.title = c("Aroma Score",""))$aroma

plot_11 <- plot_model(model_5, type = "pred", title = "Probability of Poor Quality Class",
           axis.title = c("Flavor Score", ""))$flavor

plot_12 <- plot_model(model_5, type = "pred", title = "Probability of Poor Quality Class",
           axis.title = c("Acidiry Score", ""), element_text(size=1))$acidity

grid.arrange(plot_10, plot_11, plot_12, nrow=2)
```




# Extend Analysis -- Prediction Assesment. 

## Confusion Matrix
```{r ,echo=FALSE,warning=FALSE, message=FALSE, fig.align = "center",out.width = '68%',  fig.pos = 'H'}
data$Qualityclass <- ifelse(data$Qualityclass == "Good", 0, 1)
data$Qualityclass <- as.factor(data$Qualityclass)
##set up train data and test data. 
train_sub=sample(nrow(data), 8/10*nrow(data))
train_data <- data[train_sub, ]
test_data <- data[-train_sub, ]

##check the fitting information at train_data_set. 
model_for_train_data <- glm(Qualityclass~aroma+flavor+acidity,family = binomial(link="logit"), data=train_data)
summary(model_for_train_data)

##add predicted data to the test_data_set. In which, the probability greater than 0.65 will be served as poor quality of coffee.
predict_logistic <- as.numeric(predict(model_for_train_data, 
                                       newdata=test_data,
                                       type="response")>0.65)
levels(data$Qualityclass)
conMat <- confusionMatrix(factor(predict_logistic), 
                          factor(test_data$Qualityclass), positive="1")
## Accuracy
conMat[["overall"]] %>% 
  kable(col.names = "Value",
        caption = '\\label{tab:acc} Accuracy of Prediction.') %>% 
  kable_styling(font_size = 10, latex_options="HOLD_position")
##sensitivity and specificity
conMat[["byClass"]] %>% 
  kable(col.names = "Value",
        caption = '\\label{tab:sens} The Resule of Sensitivity and Specificity of Prediction.') %>% 
  kable_styling(font_size = 10, latex_options="HOLD_position")

## Reference and prediction information
conMat[["table"]] %>% 
  kable(col.names = c("Actual Good","Actual Bad"),
        row.names = TRUE,
        caption = '\\label{tab:predic} Confuse table.',
        booktabs = TRUE, linesep = "") %>% 
  kable_styling(font_size = 10, latex_options="HOLD_position")
```





## ROC Curve 

```{r ROC curve, warning=FALSE, message=FALSE, fig.align = "center",out.width = '68%',  fig.pos = 'H',fig.cap = "\\label{fig:ROC} ROC cureve for model predicton", fig.pos = 'H'}
roc.cruve <- roc(test_data$Qualityclass,
                 predict_logistic,
                 plot=TRUE, print.thres=TRUE, print.auc=TRUE,
                 levels=c(0,1), direction="<")
```

# Conclusion 
