---
title: "Project2"
author: "Hiba Hajali"
date: "20/03/2022"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(janitor)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(gridExtra)
library(rcompanion)

```

# Introduction 

```{r}
##import data 
data <- read.csv("dataset8.csv")
```

The numbers of the missing values in each column:


```{r}

##check for missing values using sapply
sapply(data, function(x) sum(is.na(x)))

```

The data after we remove the missing values:

```{r}
##removing missing values
data <- na.omit(data)
glimpse(data)

##replace good/poor with the value 1 and 0
#data$Qualityclass <- ifelse(data$Qualityclass == "Good", 1, 0)

##assign the value to keep in variable
#Qualityclass <- data$Qualityclass
#country_of_origin <- data$country_of_origin
```


The number of unique values in country of origin: 

```{r}
##check for number of unique values in country of origin and harvest year using sapply
length(unique(data$country_of_origin))

```

The number of unique values in harvest year:
```{r}
length(unique(data$harvested))
```

# Explantory Analysis 

The correlation between the quantitative variables:

```{r plots, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE}
##correlation between variables. 
pair_cor <- data %>% 
    select(Qualityclass,
           harvested,
           altitude_mean_meters,
           aroma,
           flavor,
           acidity,
           category_two_defects) 
pair_cor$Qualityclass <- as.factor(pair_cor$Qualityclass)


#Quality class vs aroma.
plot_1  <-  ggplot(data = pair_cor, 
                   aes(x = Qualityclass, 
                       y = aroma, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality class vs aroma.", y = "Aroma") +
            theme(legend.position = "none")

ggpairs(pair_cor) +
    theme(text = element_text(size=10))

```

Box plots showing the distribution of the quantitative variables  

```{r boxplots, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE}
plot__1  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = aroma, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Aroma") +
            theme(legend.position = "none")

plot_2  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = flavor, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Flavor") +
            theme(legend.position = "none")

plot_3  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = acidity, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Acidity") +
            theme(legend.position = "none")

plot_4  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = category_two_defects, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Category Two Defects") +
            theme(legend.position = "none")

plot_5  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = altitude_mean_meters, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Altitude mean (meters)") +
            theme(legend.position = "none")

plot__1 
plot_2
plot_3 
plot_4
plot_5

```


bar charts:

```{r}
#bar chart showing the quality class in each country 
plot_6  <-  ggplot(data = data, 
                   aes(x = country_of_origin, 
                       group = Qualityclass)) +
            geom_bar(aes(y = ..prop..,fill=  Qualityclass),
                     stat = "count", position = "dodge")
plot_6

#bar chart showing the quality class in each harvest year 
plot_7  <-  ggplot(data = data, 
                   aes(x = harvested, 
                       group = Qualityclass)) +
            geom_bar(aes(y = ..prop..,fill=  Qualityclass),
                     stat = "count", position = "dodge")
plot_7

```

The percentages:

Table showing the percentage of the quality classes for each country

```{r}
#table showing the percentage of the quality classes for each country  
data %>%
    tabyl(country_of_origin, Qualityclass) %>%
    adorn_percentages() %>%
    adorn_pct_formatting() %>%
    adorn_ns() 

```

Table showing the percentage of the quality classes for each harvest year:

```{r}
#table showing the percentage of the quality classes for each harvest year   
data %>%
    tabyl(harvested, Qualityclass) %>%
    adorn_percentages() %>%
    adorn_pct_formatting() %>%
    adorn_ns()
```


## Formal Analsis 

Model 1:

$$Qualityclass \sim country\_of\_origin + aroma + flavor + acidity + category\_two\_defects + altitude\_mean\_meters$$

```{r ,warning=FALSE, message=FALSE}
#setting the quality class as factor 
data$Qualityclass <- as.factor(data$Qualityclass)

model_1 <- glm(formula = Qualityclass ~ .,family = binomial(link = "logit"),
               data = data)
summ(model_1)    

```

Model 2:

$$Qualityclass \sim country\_of\_origin + aroma + flavor + acidity + category\_two\_defects + harvested$$

```{r, warning=FALSE, message=FALSE}
model_2 <- glm(formula = Qualityclass ~ country_of_origin +aroma+flavor+acidity+category_two_defects+harvested,family = binomial(link = "logit"),
               data = data)
summ(model_2)

```


Model 3:

$$Qualityclass \sim country\_of\_origin + aroma + flavor + acidity + category\_two\_defects$$

```{r, warning=FALSE, message=FALSE}
model_3 <- glm(formula = Qualityclass ~ country_of_origin +aroma+flavor+acidity+category_two_defects,family = binomial(link = "logit"),
               data = data)
summ(model_3)
```

Model 4:

 $$Qualityclass \sim aroma + flavor + acidity + category\_two\_defects$$
 
```{r, warning=FALSE, message=FALSE}
model_4 <- glm(formula = Qualityclass ~ aroma+flavor+acidity+category_two_defects,family = binomial(link = "logit"),
               data = data)
summ(model_4)
```


Model 5:

 $$Qualityclass  \sim  aroma + flavor + acidity$$

```{r, warning=FALSE, message=FALSE}
#best model
model_5 <- glm(formula = Qualityclass ~ aroma+flavor+acidity ,family = binomial(link = "logit"),
               data = data)
summ(model_5)

```


Models comparison:

```{r, warning=FALSE, message=FALSE}
#compare 5 models
compareGLM(model_1, model_2, model_3, model_4,model_5)
```



Odds Plot:

```{r odds plot, warning=FALSE, message=FALSE}
#Odds Plot
plot_model(model_5, show.values = TRUE, transform = NULL,
           title = "Log-Odds (features)", show.p = FALSE)

```


Confidence intervals:

```{r, warning=FALSE, message=FALSE}
#Confidence intervals
confint(model_5)
```

## Conclusion 
