---
title: "Project2"
author: "Hiba Hajali"
date: "20/03/2022"
output:
  pdf_document:
    latex_engine: pdflatex
    number_sections: yes
    keep_tex: true
  html_document:
    df_print: paged
fig_caption: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(janitor)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(gridExtra)
library(rcompanion)
library(knitr)
library(kableExtra)
library(pROC)
library(caret)
library(huxtable)
```

# Introduction 

Coffee is one of the most popular beverages worldwide and has a vast market. Research on coffee quality can help coffee farmers understand the quality of the coffee they grow to make more accurate market planning. The researchers obtained data containing features of coffee and its production from the Coffee Quality Institute, a coffee research institute. They used this data to analyze the impact of these coffee features (such as acidity) on coffee quality scores. 

In the following sections, the researchers will use the Generalized Linear Model to model the Qualityclass variables, obtain the optimal model by comparison, and analyze each variable to determine its impact on coffee quality.
```{r}
##import data 
data <- read.csv("dataset8.csv")
```
# Explanatory Analysis 

The numbers of the missing values in each column. From the output we find that NA values is concentrate on variable of 'harvested' and 'altitude mean meters'.
```{r missing value checking}

##check for missing values using sapply.
sapply(data, function(x) sum(is.na(x)))
```

The data after we remove the missing values, 858 observations were obtained. And data set includes 8 variables in total. To answer the research question, the Quality class serves as response variables, the country_of_origin, aroma, flavor, acidity, category_two_defects, altitute_mean_meters and harvested serves as explanatory variables.

Since there are only two kinds of results, good and bad, the response variable is dichotomous and follows binomial distribution, which is suitable for binary logistic regression in GLM, not multivariate logistic regression.

```{r na value removal}
##removing missing values
data <- na.omit(data)
glimpse(data)
```



The number of unique values in country of origin: 
```{r Number of country}
##check for number of unique values in country of origin and harvest year using sapply
length(unique(data$country_of_origin))

```

The number of unique values in harvest year:
```{r length of variable year}
length(unique(data$harvested))
```
Correlation analysis diagram shows that there are strong positive correlations between flavor,aroma and acidity, so multicollinearity should be paid attention to in the subsequent modeling process. The correlation between other variables is weak.  


Box plots showing the distribution of the quantitative variables.  
```{r boxplots, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
plot__1  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = aroma, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Aroma") +
            theme(legend.position = "none")

plot_2  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = flavor, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Flavor") +
            theme(legend.position = "none")

plot_3  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = acidity, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Acidity") +
            theme(legend.position = "none")

plot_4  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = category_two_defects, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Category Two Defects") +
            theme(legend.position = "none")

plot_5  <-  ggplot(data = data, 
                   aes(x = Qualityclass, 
                       y = altitude_mean_meters, 
                       fill = Qualityclass)) +
            geom_boxplot() +
            labs(x = "Quality Class", y = "Altitude mean (meters)") +
            theme(legend.position = "none")



grid.arrange(plot__1,plot_2,plot_3,nrow=1)
grid.arrange(plot_4, plot_5, nrow=1)
```

The box plot shows that three variables, aroma,flavor and acidity, have more impact on good coffee quality than bad coffee quality. However, Category Tow Defects and Altitude mean (meters) have almost the same influence on coffee quality as bad, and there are some abnormal values.  


## Bar Charts:
```{r barp, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
#bar chart showing the quality class in each country 
plot_6  <-  ggplot(data = data, 
                   aes(x = ..prop.., 
                       group = Qualityclass)) +
            geom_bar(aes(y = country_of_origin,fill=  Qualityclass),
                     stat = "count", position = "dodge") + 
  theme(axis.text.y =element_text(angle=0, hjust=1), text=element_text(size=10))
#bar chart showing the quality class in each harvest year 
plot_7  <-  ggplot(data = data, 
                   aes(x = harvested, 
                       group = Qualityclass)) +
            geom_bar(aes(y = ..prop..,fill=  Qualityclass),
                     stat = "count", position = "dodge")
```

The Percentages:

Table 1 showing the percentage of the quality classes for each country.
Colombia has the most proportion of good quality coffee, while Mexico has the most bad coffee, Ethiopia only has good coffee.
```{r,echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE, fig.align = "center",fig.pos = 'H',out.width = '68%'}
#table showing the percentage of the quality classes for each country  
data %>%
    tabyl(country_of_origin, Qualityclass) %>%
    adorn_percentages() %>%
    adorn_pct_formatting() %>%
    adorn_ns() %>% 
  kable(align='c',
   caption = "The Proportion of Quality Classs in Different Country") %>%
  kable_styling( latex_options="HOLD_position")

```




Table 2 indicates the percentage of the quality classes for each harvest year. In 2012, 2016 and 2017 there was a larger proportion of bad coffee, and in all other years there was a larger proportion of good coffee.
```{r, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE, fig.align = "center",  fig.pos = 'H', out.width = '68%'}
#table showing the percentage of the quality classes for each harvest year   
data %>%
    tabyl(harvested, Qualityclass) %>%
    adorn_percentages() %>%
    adorn_pct_formatting() %>%
    adorn_ns() %>% 
  kable(align='c',
   caption = "The Proportion of Quality Classs in Different Harvested Year") %>%
  kable_styling(latex_options="HOLD_position")
```


# Formal Anaylsis 
## Multicollinearity Checking.

Before start to establish fitting model, the multicollinearity examine between our potential variables. Here, kappa function helps to evaluate the multicollinearity. kappa value for the explanatory variables of this research is 21.31, that is lower than 100. In general, it considered that the multicollinearity is weak.
```{r multiliconearility, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
multico.variables <- data %>% 
  select(aroma,
         flavor,
         acidity,
         category_two_defects,
         altitude_mean_meters,
         harvested) %>% 
  cor()

kappa(multico.variables, exact=TRUE)
```


## Models comparison and Selection:

```{r model 1,eval=TRUE, echo=FALSE, fig.align="center", fig.pos='H', message=FALSE, ,warning=FALSE, out.width='68%'}
#setting the quality class as factor 
data$Qualityclass <- as.factor(data$Qualityclass)

model_1 <- glm(formula = Qualityclass ~ .,family = binomial(link = "logit"),
               data = data)
```



```{r model 2, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
model_2 <- glm(formula = Qualityclass ~ country_of_origin +aroma+flavor+acidity+category_two_defects+harvested,family = binomial(link = "logit"),
               data = data)
```

```{r model 3, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE,fig.align = "center",fig.pos = 'H', out.width = '68%'}
model_3 <- glm(formula = Qualityclass ~ country_of_origin +aroma+flavor+acidity+category_two_defects,family = binomial(link = "logit"),
               data = data)
```

```{r model 4, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
model_4 <- glm(formula = Qualityclass ~ aroma+flavor+acidity+category_two_defects,family = binomial(link = "logit"),
               data = data)
```


```{r model 5, warning=FALSE, message=FALSE, echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE,fig.align = "center",  fig.pos = 'H',out.width = '68%'}
#best model
model_5 <- glm(formula = Qualityclass ~ aroma+flavor+acidity ,family = binomial(link = "logit"),
               data = data)

summary(model_5)
```


```{r baseline checking, echo=TRUE}
levels(data$Qualityclass)
#check the baseline and contributor for our response variables in models.  
#the first returned result is our baseline and the second returned is the contributor.
```

According to the characteristics of our response variable (Quality class), the GLM model of logistic family has been used to modelling. For model_1 to model_5, it estimates the change of the coffee bean with poor quality class according to the response baseline, which is good quality. While, this change is reflect on a log-odds scale. 



```{r model comparison, warning=FALSE, message=FALSE, fig.align = "center", fig.pos = 'H', out.height=12, out.width=6}
#compare 5 models
compareGLM(model_1, model_2, model_3, model_4,model_5) %>% 
  kable(digits=3,align='c',
   caption = "The Result of Model comparison") %>%
  kable_styling(font_size=8, latex_options="HOLD_position")
```

the p value 




The strategy to establish model is the top-down approach. Starting with a model contains all variables, then remove the variables one by one. Evaluate it by AIC BIC and p values. 



After the result of model comparison, formula of final model is given by:
$$ln\left(\frac{p_{Poor}}{1-p_{Poor}}\right) = \alpha  + \beta_1 \cdot \textrm{Aroma}+ \beta_2 \cdot \textrm{Flavor} + \beta_3 \cdot \textrm{Acidity}$$
where $p=Prob(Poor)$ and $1-p=Prob(Good)$. In which, the instructor is Poor quality of coffee.  

##  Information of final model.  
```{r final model information}
summary(model_5)
```

```{r odds plot, warning=FALSE, message=FALSE,out.width = '68%', fig.align = "center",  fig.pos = 'H'}
#Odds Plot
plot_model(model_5, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Poor instructor)", show.p = FALSE)

```

## Confidence Intervals:
```{r , warning=FALSE, message=FALSE, fig.align = "center",  fig.pos = 'H'}
#Confidence intervals
confint(model_5) %>% 
  kable(caption = '\\label{tab:CI5} Confidence Intervals for log odds in Model 5') %>% 
  kable_styling(font_size = 10,latex_options="HOLD_position")
```


## The Predictive Probability Plot with marginal effect: 
```{r probability plot,warning=FALSE, message=FALSE, fig.align = "center",  fig.pos = 'H' }
log.odds.est <- data %>%
  mutate(logodds.poor = predict(model_5))

plot_10 <- plot_model(model_5, type = "pred", title = "Probability of Poor Quality Class",
           axis.title = c("Aroma Score",""))$aroma

plot_11 <- plot_model(model_5, type = "pred", title = "Probability of Poor Quality Class",
           axis.title = c("Flavor Score", ""))$flavor

plot_12 <- plot_model(model_5, type = "pred", title = "Probability of Poor Quality Class",
           axis.title = c("Acidiry Score", ""), element_text(size=1))$acidity

grid.arrange(plot_10, plot_11, plot_12, nrow=2)
```




# Extend Analysis -- Prediction Assesment. 

## Confusion Matrix
```{r ,echo=FALSE,warning=FALSE, message=FALSE, fig.align = "center",out.width = '68%',  fig.pos = 'H'}
data$Qualityclass <- ifelse(data$Qualityclass == "Good", 0, 1)
data$Qualityclass <- as.factor(data$Qualityclass)
##set up train data and test data. 
train_sub=sample(nrow(data), 8/10*nrow(data))
train_data <- data[train_sub, ]
test_data <- data[-train_sub, ]

##check the fitting information at train_data_set. 
model_for_train_data <- glm(Qualityclass~aroma+flavor+acidity,family = binomial(link="logit"), data=train_data)
summary(model_for_train_data)

##add predicted data to the test_data_set. In which, the probability greater than 0.65 will be served as poor quality of coffee.
predict_logistic <- as.numeric(predict(model_for_train_data, 
                                       newdata=test_data,
                                       type="response")>0.65)
levels(data$Qualityclass)
conMat <- confusionMatrix(factor(predict_logistic), 
                          factor(test_data$Qualityclass), positive="1")
## Accuracy
conMat[["overall"]] %>% 
  kable(col.names = "Value",
        caption = '\\label{tab:acc} Accuracy of Prediction.') %>% 
  kable_styling(font_size = 10, latex_options="HOLD_position")
##sensitivity and specificity
conMat[["byClass"]] %>% 
  kable(col.names = "Value",
        caption = '\\label{tab:sens} The Resule of Sensitivity and Specificity of Prediction.') %>% 
  kable_styling(font_size = 10, latex_options="HOLD_position")

## Reference and prediction information
conMat[["table"]] %>% 
  kable(col.names = c("Actual Good","Actual Bad"),
        row.names = TRUE,
        caption = '\\label{tab:predic} Confuse table.',
        booktabs = TRUE, linesep = "") %>% 
  kable_styling(font_size = 10, latex_options="HOLD_position")
```





## ROC Curve 
```{r ROC curve, warning=FALSE, message=FALSE, fig.align = "center",out.width = '68%',  fig.pos = 'H',fig.cap = "\\label{fig:ROC} ROC cureve for model predicton", fig.pos = 'H'}
roc.cruve <- roc(test_data$Qualityclass,
                 predict_logistic,
                 plot=TRUE, print.thres=TRUE, print.auc=TRUE,
                 levels=c(0,1), direction="<")
```

# Conclusion 

Aroma,flavor and acidity affect the quality of coffee. With the increase of aroma, the probability of good coffee quality increases. Similarly, as flavor and acidity increase, the probability of good coffee quality increases. The variation of flavor has the greatest impact on the quality of coffee, while the variation of acidity has the least impact on the quality of coffee.

All but three of the variables in the final model had very little effect on the quality of the coffee.

Futher study: Evaluate the predictive power of the selected model by comparing the values of accuracy, sensitivity, specificity and AUC with the confusion matrix and ROC.
